---
title: "p8105_hw2_yl5227"
author: "Jennifer Li"
date: "2023-10-03"
output: github_document
---

```{r message=FALSE}
library(tidyverse)
library(readxl)
```


# Problem 1


import datasets

```{r}
df_pols = read_csv("data/pols-month.csv")

df_unemployment = read_csv("data/unemployment.csv")

df_snp = read_csv("data/snp.csv")
```


data cleaning

```{r}
df_pols =
  df_pols |> 
  separate(mon, into = c("year", "month", "day"), sep = "-") |> 
  mutate(
    month = case_match(
      month,
    "01" ~ "Jan",
    "02" ~ "Feb",
    "03" ~ "March",
    "04" ~ "April",
    "05" ~ "May",
    "06" ~ "June",
    "07" ~ "July",
    "08" ~ "Aug",
    "09" ~ "Sept",
    "10" ~ "Oct",
    "11" ~ "Nov",
    "12" ~ "Dec"
    )) |> 
    mutate(president = ifelse(prez_gop == 0, "dem", "gop")) |>
    select(-prez_dem, -prez_gop, -day)

```


```{r}
df_snp =
  df_snp |> 
    separate(date, into = c("month", "day", "year"), sep = "/") |> 
    mutate(year = as.numeric(year), 
           month = as.numeric(month),
           year = ifelse(year < 16, year + 2000, year + 1900)) |> 
    arrange(year, month) |> 
    select(year, month, close) |> 
    mutate(
      month = case_match(
      month,
    1 ~ "Jan",
    2 ~ "Feb",
    3 ~ "March",
    4 ~ "April",
    5 ~ "May",
    6 ~ "June",
    7 ~ "July",
    8 ~ "Aug",
    9 ~ "Sept",
    10 ~ "Oct",
    11 ~ "Nov",
    12 ~ "Dec"
    ))
    
```

```{r}
df_unemployment =
  df_unemployment |> 
    pivot_longer(Jan:Dec, 
                 names_to = "month",
                 values_to = "umemployment%") |> 
    janitor::clean_names()
```

combine 3 datasets.

```{r}
df_pols$year = as.numeric(unlist(pull(df_pols, year)))


table1 = full_join(df_pols, df_snp)
merge_table = full_join(table1, df_unemployment)
```


There are missing values denoted as 'NA' for `close` and `unemployment_percent` variables. The `df_pols` data gives us information about the number of republicans and democrats of senators and governors from `r min(pull(df_pols, year))` to `r max(pull(df_pols, year))`. The `df_unemployment` data gives us unemployment percentage from `r min(pull(df_unemployment, year))` to `r max(pull(df_unemployment, year))` by month. The `df_snp` data gives the closing values of the S&P stock index from `r min(pull(df_snp, year))` to `r max(pull(df_snp, year))`.



# Problem 2

Import and clean the `Mr. Trash Wheel` dataset.

```{r}
df_mtw = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Mr. Trash Wheel", 
             range = "A2:N586") |> 
  janitor::clean_names()|> 
  mutate(homes_powered = weight_tons * 500 / 30,
         trash_wheel = "mr",
         year=as.numeric(year)) 
```

The `Mr. Trash Wheel` data has `r nrow(df_mtw)` observations and `r ncol(df_mtw)` variables, ranging from years `r min(df_mtw$year)` to `r max(df_mtw$year)`. 



Import and clean the `Professor Trash Wheel` dataset.

```{r}
df_ptw = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Professor Trash Wheel", 
             range = "A2:M108") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = weight_tons * 500 / 30,
         trash_wheel = "professor")
```

The `Professor Trash Wheel` data has `r nrow(df_ptw)` observations and `r ncol(df_ptw)` variables ranging from years `r min(df_ptw$year)` to `r max(df_ptw$year)`. 


Import and clean the `Gwynnda Trash Wheel` dataset.

```{r}
df_gtw = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Gwynnda Trash Wheel", 
             range = "A2:L157") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = weight_tons * 500 / 30,
         trash_wheel = "gwynnda")
```

The `Gwynnda Trash Wheel` data has `r nrow(df_gtw)` observations and `r ncol(df_gtw)` variables, ranging from years `r min(df_gtw$year)` to `r max(df_gtw$year)`.


combine 3 datasets, and some statistics about the combined data.

```{r}
df_merge_tw =
  bind_rows(df_mtw, df_gtw, df_ptw)
```

The combined dataset has `r nrow(df_merge_tw) ` observations, each represents a distinct dumpster. There are `r ncol(df_merge_tw) ` variables in total include the dumpster number, the date collected, weight, volume, kinds of trash, the value of homes powered, and I have added one variable called "trash_wheel" indicates this data is from Mr. Trash Wheel's, Professor Trash Wheel's, or Gwynnda Trash Wheel's. This dataset tells us about the weight, volume, and type of trash collected by `Mr. Trash Wheel`, `Professor Trash Wheel`, or `Gwynnda Trash Wheel` for a given year from years `r min(df_merge_tw$year)` to `r max(df_merge_tw$year)`. It also tells us the number of household that the amount of electricity generated from collected trash can provide with an average household will use 30 kilowatts per day. 



## For available data, what was the total weight of trash collected by Professor Trash Wheel? What was the total number of cigarette butts collected by Gwynnda in July of 2021?

```{r}
total_weight = 
  df_merge_tw |> 
  select(trash_wheel, weight_tons) |> 
  filter(trash_wheel == "professor") |> 
  summarise(sum(weight_tons))
```

The total weight of trash collected by Professor Wheel is `r total_weight` tons.

```{r}
options(scipen = 999)
total_cigarette_butts = 
  df_merge_tw |> 
  select(trash_wheel, cigarette_butts, year, month) |> 
  filter(trash_wheel == "gwynnda" & year == 2021 & month == "July") |> 
  summarise(sum(cigarette_butts))
```

The total number of cigarette butts collected by Gwynnda in July of 2021 is `r total_cigarette_butts`.


# Problem 3

```{r}
df_mci_bl =
  read_csv("data/MCI_baseline.csv",
           skip=1,
           na=".") |> 
  janitor::clean_names() |> 
  mutate(sex = ifelse(sex == 1, "male", "female"),
         apoe4 = ifelse(apoe4 == 1, "carrier", "non-carrier")) |> 
  filter(current_age < age_at_onset|is.na(age_at_onset))
```

```{r}
df_mci_amyloid =
  read_csv("data/mci_amyloid.csv",
           skip=1,
           na=c("NA", "Na", "")) |> 
  janitor::clean_names() |> 
  pivot_longer(baseline:time_8,
               names_to = "time",
               values_to = "amyloid_ratio") |> 
  rename(id = study_id)
```


How many participants were recruited, and of these how many develop MCI? What is the average baseline age? What proportion of women in the study are APOE4 carriers?

```{r}
nrow(df_mci_bl)

sum(!is.na(pull(df_mci_bl, age_at_onset)))

mean(pull(df_mci_bl, current_age))

female_part = filter(df_mci_bl, sex == "female")
carrier_female = filter(df_mci_bl, apoe4 == "carrier" & sex == "female")
proportion <- nrow(carrier_female)/nrow(female_part)
proportion
```

In this study, `r nrow(df_mci_bl)` participants were recruited, and `r sum(!is.na(pull(df_mci_bl, age_at_onset)))` of them developed MCI. The average baseline age is `r mean(pull(df_mci_bl, current_age))`. Finally, `r proportion*100`% of women in the study are APOE4 carriers.

whether some participants appear in only the baseline or amyloid datasets

```{r}
common_part = intersect(df_mci_bl$id, df_mci_amyloid$id)

bl_only = anti_join(df_mci_bl, df_mci_amyloid, by = "id")

amyloid_only = 
  anti_join(df_mci_amyloid, df_mci_bl, by = "id") |> 
  distinct(id, .keep_all = TRUE)
```


Combine the demographic and biomarker datasets

```{r}
df_combined = inner_join(df_mci_bl, df_mci_amyloid, by = "id")
```

